Sharing and organizing research products as R packages

### 2020-05-22 manuscript/manuscript.Rmd
Update manuscript.Rmd

We sought for a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion]. Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) beyond datasets in a reproducible collection of materials. We identified the R package standard with modern online-based workflows as a solution that doesn't present overwhelming overhead for already busy researchers. Here, we discuss the R package standard for creating reproducible research material containers for data, analysis code, and metadata. Then, we introduce additional R packages for incorporating other research products into the reproducible container. These additional tools are easily accessible through our *vertical* package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.

@@

Broadly, there are 3 steps to including data in an R package: 1. Placing raw data in the "data-raw" directory, 2. creating an R script that processes the raw data and creates an R data object into the "data" directory, and 3. documenting the final data object.

>>>>

We sought a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion]. Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) beyond datasets in a reproducible collection of materials. We identified the R package standard with modern online-based workflows as a solution that doesn't present overwhelming overhead for already busy researchers. Here, we discuss the R package standard for creating reproducible research material containers for data, analysis code, and metadata. Then, we introduce additional R packages for incorporating other research products into the reproducible container. These additional tools are easily accessible through our *vertical* package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.

@@

Broadly, there are 3 steps to including data in an R package: 1. placing raw data in the "data-raw" directory, 2. creating an R script that processes the raw data and creates an R data object into the "data" directory, and 3. documenting the final data object.
